Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change

### 2020-12-30 Manuscript/HI_Model_ms.Rmd
updated manuscript files

## Abstract
The State of Hawai`i passed legislation to be carbon neutral by 2045, a goal that will partly depends on carbon sequestration by terrestrial ecosystems. However, there is considerable uncertainty surrounding the future direction and magnitude of the land carbon sink in the Hawaiian Islands. We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two radiative forcing scenarios (RCPs 4.5 and 8.5) and two land-use scenarios (low and high) over a 90-year timespan from 2010-2100. Collectively, terrestrial ecosystems of the Hawaiian Islands acted as a net carbon sink under low radiative forcing (RCP 4.5) for the entire 90-year simulation period, with low land-use change further enhancing carbon sink strength. In contrast, Hawaiian terrestrial ecosystems transitioned from a net sink to a net source of CO~2~ to the atmosphere under high radiative forcing (RCP 8.5), with high land-use accelerating this transition and exacerbating net carbon loss. A sensitivity test of the CO~2~ fertilization effect on plant productivity revealed it to be a major source of uncertainty in projections of ecosystem carbon balance. Reconciling this uncertainty in how net photosynthesis will respond to rising atmospheric CO~2~ will be essential to realistically constrain simulation models used to evaluate the effectiveness of ecosystem‐based climate mitigation strategies.

## Introduction
Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis and reducing the rate of increase in atmospheric CO~2~ [@friedlingstein_global_2019]. However, the strength of the land carbon sink remains highly uncertain because of its sensitivity to interannual variability in climate, land use and land cover change [@ahlstrom_dominant_2015; @prestele_current_2017]. The uncertainty surrounding estimates of terrestrial carbon sequestration complicates efforts to employ "natural carbon solutions", defined as conservation, restoration, and land management interventions aimed at enhancing carbon storage in natural biological reservoirs [@griscom_natural_2017]. A growing number of sub-national jurisdictions have pledged to incorporate natural climate solutions as part of their stated goal to achieve carbon neutrality by mid-century [].

>>>>

## Abstract
The State of Hawai`i passed legislation to be carbon neutral by 2045, a goal that will partly depend on carbon sequestration by terrestrial ecosystems. However, there is considerable uncertainty surrounding the future direction and magnitude of the land carbon sink in the Hawaiian Islands. We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two radiative forcing scenarios (RCPs 4.5 and 8.5) and two land-use scenarios (low and high) over a 90-year timespan from 2010-2100. Collectively, terrestrial ecosystems of the Hawaiian Islands acted as a net carbon sink under low radiative forcing (RCP 4.5) for the entire 90-year simulation period, with low land-use change further enhancing carbon sink strength. In contrast, Hawaiian terrestrial ecosystems transitioned from a net sink to a net source of CO~2~ to the atmosphere under high radiative forcing (RCP 8.5), with high land-use accelerating this transition and exacerbating net carbon loss. A sensitivity test of the CO~2~ fertilization effect on plant productivity revealed it to be a major source of uncertainty in projections of ecosystem carbon balance. Reconciling this uncertainty in how net photosynthesis will respond to rising atmospheric CO~2~ will be essential to realistically constrain simulation models used to evaluate the effectiveness of ecosystem‐based climate mitigation strategies.

## Introduction
Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis and reducing the rate of increase in atmospheric CO~2~ [@friedlingstein_global_2019]. There is increasing recognition among policymakers that natural and angricultural ecosystems can contribute to climate mitigation, which has given rise to the popularity of “natural carbon solutions” [@cameron_ecosystem_2017]. Natural climate solutions are defined as conservation and land management efforts aimed at enhancing carbon storage in natural biological reservoirs [@griscom_natural_2017]. A growing number of sub-national jurisdictions have set goals to achieve carbon neutrality by mid-century, and plan to incorporate natural climate solutions because they are seen as cost-effective and readily available [@averchenkova_climate_2017; @cameron_ecosystem_2017; @fargione_natural_2018]. However, effective implementation is complicated by the considerable uncertainty surrounding the future direction and magnitude of the land carbon sink, especially at the regional scale.

Interannual variability in climate, land use, and land cover change [@ahlstrom_dominant_2015; @prestele_current_2017].

Hawaii was the first U.S. state to enact legislation supporting commitments and goals of the Paris climate agreement.
